Using Scagnostics to find Interesting Projections of Multivariate Data

### 2018-09-06 scagnostics.Rmd
Commiting old changes before rewrite.

* Outlying: Outliers are defined in non-parametric fashion via the edge length in the MST, where the 25% and 75% quantil edge length define a threshold value $w=q_{75}+1.5(q_{75}-q_{25})$ above which an edge is considered long. The outlying measure is then calculated as $c_{outlying} = \frac{\mathrm{Total\ length\ of\ long\ edges}}{\mathrm{Total\ length\ of\ all\ edges}}$. Note that outlying points are not considered in the calculation of the other measures.
* Skewed: Skewness is defined via the distribution of edge length in the MST, $q_{skew}=\frac{q_{90}-q{50}}{q_{90}-q_{10}}$. This is inverted as $c_{skew}= 1-w(1-q_{skew})$ to account for binning effects that result in decreasing values of $q_{skew}$ with increasing n.
* Sparse: A second measure based on the distribution of edge lenths, we define $c_{sparse} = w q_{90}$. It takes large values if points are concentrated in well separated parts of the plane, and low values if they are distributed approximately on a lattice.
* Clumpy: To indicate the clustering of points the Clumpy measure is defined as $c_{clumpy}=\max\limits_{j}[1-\max\limits_{k}[\frac{length(e_k)}{length(e_j)}]]$ where we maximise the edge length $length(e_k)$ in the smaller of two subgraphs generated by removing a single edge $e_j$ from the MST. Note that $c_{clumpy}$ will take large values (close to one) if a long edge is separating clustered points connected by short edges.
* Striated: We define an edge as striated if they have a large angle with one of the neighboring edges, concretely $cos(angle) < -0.75$, and the corresponding measure $c_{striate} = \frac{Number of striated edges}{Number of all edges}$. FIXME is this really correct?
* Convex: The convex measure is defined as $c_{convex}= w\frac{area(A)}{area(H)}$ where $A$ is the alpha hull and $H$ the convex hull.
* Skinny: The skinnyness of a polygon can be measured as the ratio of the perimiter to the area. We define $c_{skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalisation is chosen such that $c_{skinny} = 0$ for a circle, and values close to one for skinny polygons.
* Stringy: need to check, should be $c_{stringy} = \frac{diameter(MST)}{lenght(MST)}$ where diameter is longest connected path, and lenght the total lenght (sum of all edgest), i.e. we are measuring branch structure, no branches means $c_{stringy}=1$.

@@

# Possible modifications

## Smoothing

## Jittering

## Tour optimisation

>>>>

* Outlying: Outliers are defined in non-parametric fashion via the edge length in the MST, where the 25% and 75% quantil edge length define a threshold value $w=q_{75}+1.5(q_{75}-q_{25})$ above which an edge is considered long. The outlying measure is then calculated as $c_{outlying} = \frac{\mathrm{Total\ length\ of\ long\ edges}}{\mathrm{Total\ length\ of\ all\ edges}}$. Note that outlying points are not considered in the calculation of the other measures.
* Skewed: Skewness is defined via the distribution of edge length in the MST, $q_{skew}=\frac{q_{90}-q{50}}{q_{90}-q_{10}}$. This is inverted as $c_{skew}= 1-w(1-q_{skew})$ to account for binning effects that result in decreasing values of $q_{skew}$ with increasing n.
* Sparse: A second measure based on the distribution of edge lenths, we define $c_{sparse} = w q_{90}$. It takes large values if points are concentrated in well separated parts of the plane, and low values if they are distributed approximately on a lattice.
* Clumpy: To indicate the clustering of points the Clumpy measure is defined as $c_{clumpy}=\max\limits_{j}[1-\max\limits_{k}[\frac{length(e_k)}{length(e_j)}]]$ where we maximise the edge length $length(e_k)$ in the smaller of two subgraphs generated by removing a single edge $e_j$ from the MST. Note that $c_{clumpy}$ will take large values (close to one) if a long edge is separating clustered points connected by short edges.
* Striated: We define an edge as striated if they have a large angle with one of the neighboring edges, concretely $cos(angle) < -0.75$, and the corresponding measure $c_{striate} = \frac{Number of striated edges}{Number of all edges}$. FIXME is this really correct? Maybe reason for discreet behaviour observed below..
* Convex: The convex measure is defined as $c_{convex}= w\frac{area(A)}{area(H)}$ where $A$ is the alpha hull and $H$ the convex hull.
* Skinny: The skinnyness of a polygon can be measured as the ratio of the perimiter to the area. We define $c_{skinny} = 1 -  \frac{\sqrt{4\pi area(A)}}{perimeter(A)}$, where the normalisation is chosen such that $c_{skinny} = 0$ for a circle, and values close to one for skinny polygons.
* Stringy: need to check, should be $c_{stringy} = \frac{diameter(MST)}{lenght(MST)}$ where diameter is longest connected path, and lenght the total lenght (sum of all edgest), i.e. we are measuring branch structure, no branches means $c_{stringy}=1$.

@@

# Possible modifications
We next study possible modifications that will enable the use of scagnostics as an index for guided tours.
First, given the often large variation in scagnositcs measures between neighbouring projections, we consider smoothing the calculation, either via averaging over $n$ previous projections, or by jittering either the points or the projection matrix and averaging the resulting scagnostics measures over jittered projections.
We then consider how different optimizers may be more adapt for an optimization over the erratic scagnostics measures.

## Smoothing
First we consider smoothing the scagnostics index by averaging over $n$ previous projections, the resulting behaviour is shown in Figure...

## Jittering
Another method of smoothing the scagnostics index would consist in jittering either the projected data points, or the projection itself.
We show the impact of both methods in Figure ...

## Tour optimisation
How does current optimizer work? What is the advantege of the alternative optimizer? Show results we get using each optimizer
